Machine Learning, Building Vintage and Property Values

### 2019-09-04 text/manuscript_assa.tex
clean slate

---
title: "Machine Learning, Building Vintage and Property Values"
author: "Thies Lindenthal* (University of Cambridge) & Erik B. Johnson (University of Alabama)"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: Externalities of architecture, machine learning
fig_caption: yes
header-includes:
   - \usepackage{soul}
   - \usepackage{color}
   - \sethlcolor{yellow}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{graphicx}
   - \usepackage{grffile}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{multirow}
   - \doublespacing
   - \usepackage[backend=bibtex, style=authoryear]{biblatex}
fig_width: 6
fig_height: 4
output: 
  pdf_document:
    keep_tex: true
bibliography: references.bib
abstract: "This paper first introduces an algorithm that collects pictures of individual buildings from Google Street View. Second, it trains a deep convolutional neural network (CNN) to classify residential buildings into architectural styles, taking into account spatial dependencies of building vintages. Third, it investigates whether architectural styles influence house prices. For re-sales, the architectural style is a significant determinant of transaction prices while no such effect is found for new buildings. This indicates that any premia are for vintage-related quality characteristics and not for a home's ``beauty''." 
---

# Introduction
\let\svthefootnote\thefootnote
\let\thefootnote\relax\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp and seminar participants at UCL/Bartlett and at the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}
\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote
If new developments were only pleasing to the eye, so Britain's *Building Better, Building Beautiful Commission*, nimbyism would cease and housing supply could finally reach the levels demanded by a growing and more affluent population [@economistlineofbeauty2018]. The inaugural chair of this new governmental commission boldly suggests that we should *``build, as our Georgian and Victorian forebears built* [...]. *All objections to new building would slip away in the sheer relief of the public''* [@scruton2018]. Even Prince Charles, in similar spirit, put forward ten principles for urban growth and architecture that emphasize tradition and aesthetics [@princecharles2014]. When governments and princes occupy themselves with beauty, economists surely may take a closer look as well. In this paper, we aim to identify the architectural styles of residential buildings using computer vision techniques and to empirically search for any transaction price premia associated with these styles. If aesthetics were as strong a force in the built environment as claimed, we should indeed find a significant effect on property prices.

This paper will assess architectural aesthetics heads on and analyze the visual appearance of residential buildings using street level images. Recent work illustrates how street level imagery and machine learning classification is an efficient and powerful combination for measuring previously unobserved characteristics of the urban environment. @Naik2017 describe how neighborhood demographics may impact the physical appearance of neighborhoods. @Gebru2017 use classified vehicle make and model to predict income, race, education, and voting patterns at the precinct level. @Glaeser2016 predict income in New York City. @Naik2016AER create a neighborhood safety based Streetscore which is shown to be highly correlated with neighborhood population density and household income. @Nadai2016 find that greenery and street facing windows contribute to a positive appearance of safety while @Liu2017 evaluate the quality and upkeep of the built environment along Beijing's streets.

@GlaeserKincaidNaik2018 push the level of observation from the block, street, or street-section level to the individual \emph{building level}. Utilizing images of buildings' exteriors collected from Google Street View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}}, and to a lesser degree interior images from Zillow, they find that looks matter, at least in Boston: A one standard deviation improvement of a building's exterior is associated with an additional USD 70,000 in home value. Intuitively, the link between good looks and value is bi-directional: The appearance of buildings that went through foreclosure deteriorated significantly [@GlaeserKincaidNaik2018].

Zooming in at individual buildings significantly increases the benefits of using mass collected street level imagery in economic research: property characteristics previously deemed "unobservable" can be directly observed in an accurate, objective, automatic and therefore cost effective way. Deriving additional variables from unconventional data sources like 3D airborne laser scanning (or in our case Google Street View) is fundamental since this provides "essential determinants influencing real estate prices [which] are constantly missing and are not accessible in official and mass appraiser databases" [@Helbich2013].

@@

## Building vintage and property values

The aesthetics of different architectural eras make for an easy cocktail party conversation topic -- and they also influence market prices. Using architectural assessments by human experts, @Buitelaar2017 find a sizable premium of 5\% for new buildings in the Netherlands that refer to traditional styles and a staggering 15\% premium for new buildings that closely follow traditional shapes, facade composition and details. Their study carefully disentangles the architectural style from other unobserved characteristics such as building quality, differences in location or year of construction. These controls are crucial, as earlier work has established that age and vintage variables tend to be highly correlated. @Coulson2008, for instance, suggest a non-parametric estimator and establish a U-shaped age function and distinct price discounts for postwar and contemporary vintages (vis-à-vis more historic vintages). @Francke2017a investigate the depreciation of residential real estate in the Netherlands and decompose land versus structure values singling out the effect of "physical deterioration, functional obsolescence and and vintage effects". They find that buildings from the 1930s carry a strong price premium.

A large scale assessment of buildings' exteriors will allow for an analysis of the externalities of architecture. Buildings hardly ever stand in isolation and @Ahlfeldt2012 investigate the influence a building's architecture exerts on its surroundings. They observe a positive price effect for residential buildings in the direct proximity of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A building's exterior does not need to be an architectural masterpiece to co-determine the value of other houses close by. Homogeneity of building shapes within street segments does influence property values. A similarly shaped neighboring building is value enhancing while proximity to a wildly different neighboring shape, everything else remaining equal, is detrimental to property values [@Lindenthal2017b].

Unfortunately, the traditional approaches chosen by e.g. @Buitelaar2017 or @Ahlfeldt2012 do not scale well. Each observation has to be classified into architectural styles by a human expert, which is time-consuming and costly, imposing an upper limit on the number of observations and the level of detail captured for each observation in any given sample. Google Street View offers a solution to this nexus as it captures the images of almost all buildings in many cities around the world at high level of accuracy and detail. The challenge we address in this paper is to extract and utilize building level information from this ubiquitous sensor.

# Methodology

## Image collection

@@

Previous work sourcing imagery from Google Street View has mostly focused on neighborhood or precinct characteristics where exact spatial assignment of objects is not required and street-level images suffice. For major US cities, the accuracy of building image search results is higher, which should help studies such as @GlaeserKincaidNaik2018.

@@

By combining land registry maps with the snapshot location, we are able to automatically build a set of building frontage images for approximately 48,000 properties in Cambridge (UK). The Ordnance Survey maps do not provide any use type classifications so we collect images on all buildings that have a ground plate between 40 and 250 $m^2$. This method, however, could easily be adapted to photograph buildings in locations for which high-resolution building level maps are not available by using LIDAR based building outlines instead, which would be in the spirit of @Glaeser2016.

### From pixels to computer vision

We download all pictures at the highest resolution offered by Google's Street View API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}} (640x640 color pixels). We then obtain 2048-dimensional feature vectors for each picture, using the \emph{Inception-v3} deep convolutional neural network [@Szegedy2015]. \emph{Inception-v3} has been trained for the ImageNet Large Scale Visual Recognition Challenge (ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which evaluates image classification and object detection algorithms for a wide range of objects. The pretrained classifications would allow us to identify pets, vehicles or people on the pictures -- assessing architectural finesse, however, is beyond the canned classifiers' capabilities.

@GlaeserKincaidNaik2018 rely on an different ILSVRC competitor, \emph{Resnet-101} [@He2016]. They reduce the extracted feature vectors to lower dimensionality (1024 to 100 dimensions) based on principal component analysis (PCA). We follow a different strategy and double the dimensionality by including the feature vector of the building's nearest neighbor. Doubling up allows us to model spatial dependencies in building styles, similar to spatially correlated land cover classifications in @Ghimire2010a.

## Image classification

@@

* \emph{Georgian} (c1714--1837) houses feature key characteristics such as sash windows, fan lights above doors, the use of stucco on facades, often wrought work grilles, railings etc.\label{def_eras}

* In the \emph{Early Victorian} era (c1837--c1870s), a growing taste for individualized embellishment led to the development of elaborate features such as carved barge boards or finials. The development of sheet glass led to sash windows becoming more affordable, and, increasingly, wider.

* In the \emph{Late Victorian} era (c1870s--1901), bay windows became more and more widespread, and increasingly substantial. Stylistic movements such as the Queen Anne revival style contributed richly ornamental details to the formal repertoire employed by designers. Stained glass became more popular. \emph{Edwardian} architecture (1901-1910) tends to be less ornate than late Victorian architecture.

* The \emph{Interwar} period (1918--1939) saw the cost of building construction fall, amidst a drive to provide better housing for the working classes. New housing types were being favored.

* The \emph{Postwar} (1950-1980) era continued on this path, with an embrace of high-rise as well as low rise housing. Facades vary greatly between brick, tiling, pebbledash and render.

* The cut-off year for our \emph{Contemporary} era to begin is 1980. \emph{Revival} are contemporary buildings trying to emulate historic, mostly Victorian, architecture.

@@

We create stratified training samples of 600 buildings from each category but *Georgian*, which is the least common style in Cambridge and for which we can only sample 330 examples. This leaves a sufficiently sized out-of-sample verification data set.

@@

* an input layer the size of the feature vectors,

* one dense layer (relu) half the size of the input layer, one subsequent dropout (rate 0.5) layer,

* one dense layer (relu) a quarter the size of the input layer, one subsequent dropout (rate 0.5) layer,

* and the final dense output layer with softmax activation.

@@

We tabulate confusion matrices and evaluate the classification performance in terms of recall, precision and $F_1$-scores. The next step regresses the true vintage observed, captured by binomial variables $D_{TrueVint}$, against a vector of hedonic variables $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's $\mathbf{Vintage}$, estimated both with and without spatial dependencies. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients, and $\epsilon$ is the error term.

@@

The logit regressions are estimated by generalized least squares. For a well-performing classifier, the $\delta$ coefficients will be statistically significant and, more importantly, differences in the Akaike Information Criterion (AIC) allow for comparisons across the suggested ML classifiers.

Will the resulting classifiers be biased with respect to property values, picking up correlated but irrelevant cues such as greenery, upkeep, or cars brands instead of building exteriors? Are, for instance, more valuable houses more likely to be classified as the popular *Late Victorian* vintage while starter homes are disproportionally more likely to be classified as e.g. *Early Victorian*? We expect that in cases in which the ensemble provided a high confidence score, building features are clearly detectable on the pictures. For close calls, however, where models within the ensemble cannot agree, other factors might tilt the scale towards an incorrect class. A sizable difference in estimated marginal prices for estimated vintages versus values for ground truth vintages in the subsequent hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

## Building era and property price

Merging the estimated vintage classification with sales data [@LandRegistry2016a] we estimate a hedonic regression equation that establishes marginal prices for the building vintage [similar to @Moorhouse1994; @Asabere1989; @Vandell1989; @Fuerst2011; @Plaut2006], among other characteristics:
\begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i + \eta\mathbf{VintNeigh}_i + \iota\mathbf{Vintage\cdot VintNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear combination of hedonic attributes described in vector $\mathbf{X}$, vectors of year $\mathbf{Y}$ and neighborhood $\bm{Loc}$ dummy variables and the building's estimated $\mathbf{Vintage}$ and the prevailing vintage of other buildings in the direct proximity $(\mathbf{VintNeigh})$. The intercept is denoted by $\alpha$ while $\bm{\beta}$, $\bm{\delta}$, $\bm{\eta}$, $\bm{\iota}$, $\bm{\gamma}$ and $\bm{\lambda}$ are vectors of regression coefficients. $\epsilon$ is the IID error term. Heteroscedastic robust standard errors will be reported.

Are buildings from different eras imperfect substitutes catering to multiple groups of households with distinct vintage preferences? @Buitelaar2017 indicate that any premium for an architectural style must stem from either differences in construction prices (which they do not find in their Dutch sample) or from supply constraints, as new construction potentially does not capture the demand for traditional styles. For Cambridge, new supply will inevitably be of either contemporary or revival style as historic vintage buildings, by definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a subset of newly constructed buildings will show whether construction prices or supply constraints for new homes built according to different architectural styles persist -- if too few vernacular buildings were built, prices should reflect such a shortage.

# Data

Residential real estate transactions are public data in the UK, collected and published by the Land Registry [@LandRegistry2016a]. The records include the date of transaction, price paid, street address, a classification of the property type (flat, detached, semi-detached, or terraced house), the estate type (freehold or leasehold) and an indicator for newly built properties. We select transactions from Cambridge proper which were recorded between January 1995 and October 2018, excluding any leaseholds, flats and properties classified as type "other", and sales with prices below £50,000 or in excess of £2,000,000. Linking recently taken images to sales data from 24 years assumes that houses have not changed their architectural styles -- which seems reasonable. We are not aware of any conversions of just the exterior of buildings. Full redevelopments, did occur and we exclude sales for a given address pre-dating any redevelopment. Table \ref{tab:sumstats} presents summary statistics for the sample.

The Ordnance Survey *AddressBase* [@OrdnanceSurvey2017] links street addresses to building outlines on Ordnance Survey maps [@OrdnanceSurvey2017a], which allows us to calculate the buildings floor plate (in $m^2$) and to estimate the building's volume from digital elevation models [@EnvironmentAgency2015], as suggested by @Lindenthal2017. We control for the location of each building by calculating the distance to the city center proxied by Great St. Mary's Church, and non-parametrically by using 69 indicator variables for each of the smallest census tracts (*Lower Super Output Areas*, LSOA) subdividing Cambridge [@ONS2017]. LSOAs typically have 1,000--3,000 residents and 400--1,200 households of comparable economic and socio-demographic characteristics [@ONS2017b].

Overall, we collect pictures of more than 48,000 distinct buildings from the Google Street View API. The subset of 25,000 pictures that has been classified by trained architects into the seven vintages will serve as \emph{ground truth} for our study. Two additional categories identify buildings hidden by *greenery* or uninformative pictures featuring *plain walls*. The classified images are available at the authors' websites.

# Results

Almost 16,000 of the 25,000 buildings for which we have the architects' ground truth could be matched to the 23,768 sales transactions that have been recorded for Cambridge between 1995 and 2018. Thus, we observe both the machine *estimated vintage* and architect defined *ground truth* classification for the majority of buildings in the data set.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Machine Learning, Building Vintage and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Machine Learning, Building Vintage and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{04 September 2019}

\usepackage{soul}
\usepackage{color}
\sethlcolor{yellow}
\usepackage{float}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{multirow}
\doublespacing
\usepackage[backend=bibtex, style=authoryear]{biblatex}

\begin{document}
\maketitle
\begin{abstract}
This paper first introduces an algorithm that collects pictures of
individual buildings from Google Street View. Second, it trains a deep
convolutional neural network (CNN) to classify residential buildings
into architectural styles, taking into account spatial dependencies of
building vintages. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. This indicates that any premia are for
vintage-related quality characteristics and not for a home's ``beauty''.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\let\svthefootnote\thefootnote
\let\thefootnote\relax

\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp and seminar participants at UCL/Bartlett and at the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}

\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission}, nimbyism would
cease and housing supply could finally reach the levels demanded by a
growing and more affluent population (The Economist 2018). The inaugural
chair of this new governmental commission boldly suggests that we should
\emph{``build, as our Georgian and Victorian forebears built}
{[}\ldots{}{]}. \emph{All objections to new building would slip away in
the sheer relief of the public''} (Scruton 2018). Even Prince Charles,
in similar spirit, put forward ten principles for urban growth and
architecture that emphasize tradition and aesthetics (HRH The Prince of
Wales 2014). When governments and princes occupy themselves with beauty,
economists surely may take a closer look as well. In this paper, we aim
to identify the architectural styles of residential buildings using
computer vision techniques and to empirically search for any transaction
price premia associated with these styles. If aesthetics were as strong
a force in the built environment as claimed, we should indeed find a
significant effect on property prices.

This paper will assess architectural aesthetics heads on and analyze the
visual appearance of residential buildings using street level images.
Recent work illustrates how street level imagery and machine learning
classification is an efficient and powerful combination for measuring
previously unobserved characteristics of the urban environment. Naik et
al. (2017) describe how neighborhood demographics may impact the
physical appearance of neighborhoods. Gebru et al. (2017) use classified
vehicle make and model to predict income, race, education, and voting
patterns at the precinct level. Glaeser et al. (2018) predict income in
New York City. Naik, Raskar, and Hidalgo (2016) create a neighborhood
safety based Streetscore which is shown to be highly correlated with
neighborhood population density and household income. De Nadai et al.
(2016) find that greenery and street facing windows contribute to a
positive appearance of safety while Liu et al. (2017) evaluate the
quality and upkeep of the built environment along Beijing's streets.

Glaeser, Kincaid, and Naik (2018) push the level of observation from the
block, street, or street-section level to the individual
\emph{building level}. Utilizing images of buildings' exteriors
collected from Google Street
View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}},
and to a lesser degree interior images from Zillow, they find that looks
matter, at least in Boston: A one standard deviation improvement of a
building's exterior is associated with an additional USD 70,000 in home
value. Intuitively, the link between good looks and value is
bi-directional: The appearance of buildings that went through
foreclosure deteriorated significantly (Glaeser, Kincaid, and Naik
2018).

Zooming in at individual buildings significantly increases the benefits
of using mass collected street level imagery in economic research:
property characteristics previously deemed ``unobservable'' can be
directly observed in an accurate, objective, automatic and therefore
cost effective way. Deriving additional variables from unconventional
data sources like 3D airborne laser scanning (or in our case Google
Street View) is fundamental since this provides ``essential determinants
influencing real estate prices {[}which{]} are constantly missing and
are not accessible in official and mass appraiser databases'' (Helbich
et al. 2013).

@@

\hypertarget{building-vintage-and-property-values}{%
\subsection{Building vintage and property
values}\label{building-vintage-and-property-values}}

The aesthetics of different architectural eras make for an easy cocktail
party conversation topic -- and they also influence market prices. Using
architectural assessments by human experts, Buitelaar and Schilder
(2017) find a sizable premium of 5\% for new buildings in the
Netherlands that refer to traditional styles and a staggering 15\%
premium for new buildings that closely follow traditional shapes, facade
composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location or year of construction. These
controls are crucial, as earlier work has established that age and
vintage variables tend to be highly correlated. Coulson and McMillen
(2008), for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary vintages (vis-à-vis more historic vintages). Francke and
Minne (2017) investigate the depreciation of residential real estate in
the Netherlands and decompose land versus structure values singling out
the effect of ``physical deterioration, functional obsolescence and and
vintage effects''. They find that buildings from the 1930s carry a
strong price premium.

A large scale assessment of buildings' exteriors will allow for an
analysis of the externalities of architecture. Buildings hardly ever
stand in isolation and Ahlfeldt and Mastro (2012) investigate the
influence a building's architecture exerts on its surroundings. They
observe a positive price effect for residential buildings in the direct
proximity of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A
building's exterior does not need to be an architectural masterpiece to
co-determine the value of other houses close by. Homogeneity of building
shapes within street segments does influence property values. A
similarly shaped neighboring building is value enhancing while proximity
to a wildly different neighboring shape, everything else remaining
equal, is detrimental to property values (Lindenthal 2017a).

Unfortunately, the traditional approaches chosen by e.g. Buitelaar and
Schilder (2017) or Ahlfeldt and Mastro (2012) do not scale well. Each
observation has to be classified into architectural styles by a human
expert, which is time-consuming and costly, imposing an upper limit on
the number of observations and the level of detail captured for each
observation in any given sample. Google Street View offers a solution to
this nexus as it captures the images of almost all buildings in many
cities around the world at high level of accuracy and detail. The
challenge we address in this paper is to extract and utilize building
level information from this ubiquitous sensor.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{image-collection}{%
\subsection{Image collection}\label{image-collection}}

@@

Previous work sourcing imagery from Google Street View has mostly
focused on neighborhood or precinct characteristics where exact spatial
assignment of objects is not required and street-level images suffice.
For major US cities, the accuracy of building image search results is
higher, which should help studies such as Glaeser, Kincaid, and Naik
(2018).

@@

By combining land registry maps with the snapshot location, we are able
to automatically build a set of building frontage images for
approximately 48,000 properties in Cambridge (UK). The Ordnance Survey
maps do not provide any use type classifications so we collect images on
all buildings that have a ground plate between 40 and 250 \(m^2\). This
method, however, could easily be adapted to photograph buildings in
locations for which high-resolution building level maps are not
available by using LIDAR based building outlines instead, which would be
in the spirit of Glaeser et al. (2018).

\hypertarget{from-pixels-to-computer-vision}{%
\subsubsection{From pixels to computer
vision}\label{from-pixels-to-computer-vision}}

We download all pictures at the highest resolution offered by Google's
Street View
API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}}
(640x640 color pixels). We then obtain 2048-dimensional feature vectors
for each picture, using the \emph{Inception-v3} deep convolutional
neural network (Szegedy et al. 2015). \emph{Inception-v3} has been
trained for the ImageNet Large Scale Visual Recognition Challenge
(ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which
evaluates image classification and object detection algorithms for a
wide range of objects. The pretrained classifications would allow us to
identify pets, vehicles or people on the pictures -- assessing
architectural finesse, however, is beyond the canned classifiers'
capabilities.

Glaeser, Kincaid, and Naik (2018) rely on an different ILSVRC
competitor, \emph{Resnet-101} (He et al. 2016). They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the building's nearest neighbor. Doubling up allows us
to model spatial dependencies in building styles, similar to spatially
correlated land cover classifications in Ghimire, Rogan, and Miller
(2010).

\hypertarget{image-classification}{%
\subsection{Image classification}\label{image-classification}}

@@

\begin{itemize}
\item
  \emph{Georgian} (c1714--1837) houses feature key characteristics such
  as sash windows, fan lights above doors, the use of stucco on facades,
  often wrought work grilles, railings etc.\label{def_eras}
\item
  In the \emph{Early Victorian} era (c1837--c1870s), a growing taste for
  individualized embellishment led to the development of elaborate
  features such as carved barge boards or finials. The development of
  sheet glass led to sash windows becoming more affordable, and,
  increasingly, wider.
\item
  In the \emph{Late Victorian} era (c1870s--1901), bay windows became
  more and more widespread, and increasingly substantial. Stylistic
  movements such as the Queen Anne revival style contributed richly
  ornamental details to the formal repertoire employed by designers.
  Stained glass became more popular. \emph{Edwardian} architecture
  (1901-1910) tends to be less ornate than late Victorian architecture.
\item
  The \emph{Interwar} period (1918--1939) saw the cost of building
  construction fall, amidst a drive to provide better housing for the
  working classes. New housing types were being favored.
\item
  The \emph{Postwar} (1950-1980) era continued on this path, with an
  embrace of high-rise as well as low rise housing. Facades vary greatly
  between brick, tiling, pebbledash and render.
\item
  The cut-off year for our \emph{Contemporary} era to begin is 1980.
  \emph{Revival} are contemporary buildings trying to emulate historic,
  mostly Victorian, architecture.
\end{itemize}

@@

We create stratified training samples of 600 buildings from each
category but \emph{Georgian}, which is the least common style in
Cambridge and for which we can only sample 330 examples. This leaves a
sufficiently sized out-of-sample verification data set.

@@

\begin{itemize}
\item
  an input layer the size of the feature vectors,
\item
  one dense layer (relu) half the size of the input layer, one
  subsequent dropout (rate 0.5) layer,
\item
  one dense layer (relu) a quarter the size of the input layer, one
  subsequent dropout (rate 0.5) layer,
\item
  and the final dense output layer with softmax activation.
\end{itemize}

@@

We tabulate confusion matrices and evaluate the classification
performance in terms of recall, precision and \(F_1\)-scores. The next
step regresses the true vintage observed, captured by binomial variables
\(D_{TrueVint}\), against a vector of hedonic variables \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's \(\mathbf{Vintage}\), estimated both with
and without spatial dependencies. The intercept is denoted by \(\alpha\)
while \(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\gamma}\) and
\(\bm{\lambda}\) are vectors of regression coefficients, and
\(\epsilon\) is the error term.

@@

The logit regressions are estimated by generalized least squares. For a
well-performing classifier, the \(\delta\) coefficients will be
statistically significant and, more importantly, differences in the
Akaike Information Criterion (AIC) allow for comparisons across the
suggested ML classifiers.

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} vintage while starter homes are disproportionally
more likely to be classified as e.g. \emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for estimated vintages versus
values for ground truth vintages in the subsequent hedonic regression
(Eq. \ref{eq:hed}, below) would indicate such a bias.

\hypertarget{building-era-and-property-price}{%
\subsection{Building era and property
price}\label{building-era-and-property-price}}

Merging the estimated vintage classification with sales data (Land
Registry 2017) we estimate a hedonic regression equation that
establishes marginal prices for the building vintage (similar to
Moorhouse and Smith 1994; Asabere, Hachey, and Grubaugh 1989; Vandell
and Lane 1989; Fuerst, McAllister, and Murray 2011; Plaut and Uzulena
2006), among other characteristics: \begin{equation} \label{eq:hed}
  ln(Price_i) = \alpha + \beta\mathbf{X_i} + \delta\mathbf{Vintage}_i + \eta\mathbf{VintNeigh}_i + \iota\mathbf{Vintage\cdot VintNeigh}_i  +\gamma\mathbf{Y}_i + \lambda\mathbf{Loc}_i +  \epsilon_i
\end{equation}

Here, the natural logarithm of sales prices is explained by a linear
combination of hedonic attributes described in vector \(\mathbf{X}\),
vectors of year \(\mathbf{Y}\) and neighborhood \(\bm{Loc}\) dummy
variables and the building's estimated \(\mathbf{Vintage}\) and the
prevailing vintage of other buildings in the direct proximity
\((\mathbf{VintNeigh})\). The intercept is denoted by \(\alpha\) while
\(\bm{\beta}\), \(\bm{\delta}\), \(\bm{\eta}\), \(\bm{\iota}\),
\(\bm{\gamma}\) and \(\bm{\lambda}\) are vectors of regression
coefficients. \(\epsilon\) is the IID error term. Heteroscedastic robust
standard errors will be reported.

Are buildings from different eras imperfect substitutes catering to
multiple groups of households with distinct vintage preferences?
Buitelaar and Schilder (2017) indicate that any premium for an
architectural style must stem from either differences in construction
prices (which they do not find in their Dutch sample) or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. For Cambridge, new supply will inevitably be of
either contemporary or revival style as historic vintage buildings, by
definition, are not supplied any more. Estimating Eq. \ref{eq:hed} for a
subset of newly constructed buildings will show whether construction
prices or supply constraints for new homes built according to different
architectural styles persist -- if too few vernacular buildings were
built, prices should reflect such a shortage.

\hypertarget{data}{%
\section{Data}\label{data}}

Residential real estate transactions are public data in the UK,
collected and published by the Land Registry (Land Registry 2017). The
records include the date of transaction, price paid, street address, a
classification of the property type (flat, detached, semi-detached, or
terraced house), the estate type (freehold or leasehold) and an
indicator for newly built properties. We select transactions from
Cambridge proper which were recorded between January 1995 and October
2018, excluding any leaseholds, flats and properties classified as type
``other'', and sales with prices below £50,000 or in excess of
£2,000,000. Linking recently taken images to sales data from 24 years
assumes that houses have not changed their architectural styles -- which
seems reasonable. We are not aware of any conversions of just the
exterior of buildings. Full redevelopments, did occur and we exclude
sales for a given address pre-dating any redevelopment. Table
\ref{tab:sumstats} presents summary statistics for the sample.

The Ordnance Survey \emph{AddressBase} (Ordnance Survey 2017a) links
street addresses to building outlines on Ordnance Survey maps (Ordnance
Survey 2017b), which allows us to calculate the buildings floor plate
(in \(m^2\)) and to estimate the building's volume from digital
elevation models (Environment Agency 2015), as suggested by Lindenthal
(2017b). We control for the location of each building by calculating the
distance to the city center proxied by Great St.~Mary's Church, and
non-parametrically by using 69 indicator variables for each of the
smallest census tracts (\emph{Lower Super Output Areas}, LSOA)
subdividing Cambridge (Office for National Statistics 2019). LSOAs
typically have 1,000--3,000 residents and 400--1,200 households of
comparable economic and socio-demographic characteristics (Office for
National Statistics 2017).

Overall, we collect pictures of more than 48,000 distinct buildings from
the Google Street View API. The subset of 25,000 pictures that has been
classified by trained architects into the seven vintages will serve as
\emph{ground truth} for our study. Two additional categories identify
buildings hidden by \emph{greenery} or uninformative pictures featuring
\emph{plain walls}. The classified images are available at the authors'
websites.

\hypertarget{results}{%
\section{Results}\label{results}}
